\documentclass[../ai.tex]{subfiles}

\begin{document}

Satisfiability is finding whether there exists at least one satisfying model. 

Satisfiability is the dual of validity. $\psi$ is satisfiable if and only if $\neg \psi$ is not valid.

\textit{ Additionally, $KB \models \Phi$ if and only if $KB \rightarrow \Phi$ is valid if and only if $\neg (KB \rightarrow \Phi) \equiv KB \wedge \neg \Phi$ is not satisfiable. }

\textit{Satisfiability is important for other problems. Many other problems can be elegantly reduced to it.}
\\

The David-Putnam SAT-solving algorithm has a number of improvements over the previous algorithms.

We're given $KB \wedge \neg \Phi$ and need to decide whether it is not satisfiable, hence, $KB \models \Phi$ .

Suppose, we generate truth assignments as before:

\begin{itemize}
    \item If the partially built assignment already evaluates a clause $C$ to be true, we can delete $C$ and proceed.
    \item If the partially built assignment already has some clause $C$ false, we can stop extending this assignment as it can never be true.
\end{itemize}

We can extend these observations to be heuristics:

\begin{itemize}
    \item \textbf{Pure Symbol} -- A symbol that always appears positively or negatively in every clause. We can set all instances of these to true (or false respectively)
    \item \textbf{Unit Clause} -- A clause containing only one literal. We can immediately force this clause to be true, which in turn can make other clauses true.
\end{itemize}

We repeat these heuristics until we have no more unit clauses or pure symbols. Then, we continue as we did before.

Applying these heuristics can reveal pure symbols or unit clauses that were not present previously. This is known as propagation.

We can use this satisfiability solving algorithm in Wumpus World to find if $KB \models \Phi$ by testing if $KB \wedge \neg \Phi$ is not satisfiable. If it is not satisfiable, it follows that $KB \wedge \neg \Phi$ is never true and so $KB \models \Phi$ .
\\

\begin{center}
    fin.
\end{center}

\end{document}
